---
title: "p8105_hw2_"
author: "Yijia Jiang"
date: "2022-09-27"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Load the library,message=FALSE}
library(tidyverse)
```



## Problem 1 (NYC transit dataset)
```{r,warning = FALSE,message=FALSE}
# Tidy the NYC transit dataset
nyc_transit_df = read.csv("./p8105_hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude, starts_with("route"), 
         entry, vending, entrance_type, ada) %>% 
  mutate(entry = as.logical(recode(entry, "YES" = 'TRUE', "NO" = "FALSE")))


# Summary of dataset
skimr::skim(nyc_transit_df)


# Distinct stations
station = distinct(nyc_transit_df, station_name, line, .keep_all = TRUE)


# ADA compliant stations
station_ada = station %>% filter(ada == "TRUE")


# Proportion of station entrances / exits without vending allow entrance
proportion = nrow(filter(nyc_transit_df, vending == "NO", entry == "TRUE"))/nrow(filter(nyc_transit_df, vending == "NO"))


```

* The size of the resulting dataset is `r nrow(nyc_transit_df)` rows x `r ncol(nyc_transit_df)` columns.
* The names of the variables are `r names(nyc_transit_df)`.
* There are `r station %>% nrow()` distinct stations, of which `r station_ada %>% nrow()` are ADA compliant.
* The proportion of station entrances / exits without vending allow entrance is `r paste(round(100*proportion, 2), "%", sep="")`.

&nbsp;


```{r,warning = FALSE,message=FALSE}
# Reformat data so that route number and route name are distinct variables
nyc_transit_df1 = nyc_transit_df %>%
  mutate(route8 = as.character(route8),
         route9 = as.character(route9),
         route10 = as.character(route10),
         route11 = as.character(route11)) %>%
  pivot_longer(route1:route11, names_to = "route_number", values_to = "route_name")

station1 = distinct(nyc_transit_df1, station_name, line, .keep_all = TRUE)

# Distinct stations serve the A train 
station_A = filter(station1, route_name == "A")

# Stations serve the A train and ADA compliant
station_A_ada = filter(station1, route_name == "A", ada == "TRUE")
```

* `r station_A %>% nrow()` distinct stations serving the A train.
* Of the stations that serve the A train, `r station_A_ada %>% nrow()` stations are ADA compliant.

&nbsp;



## Problem 2 (Mr. Trash Wheel dataset)
```{r,warning = FALSE,message=FALSE}
# Tidy the Mr. Trash Wheel dataset
mr_trash_df= readxl::read_excel("./p8105_hw2_data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                                 sheet = "Mr. Trash Wheel",
                                 range = cellranger::cell_cols("A:N")) %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  filter(dumpster != "Grand Total") %>%
  mutate(sports_balls = as.integer(round(sports_balls, 0)),
         month = recode(month,"Decemeber" = "December"),
         dumpster = as.integer(dumpster)) %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards)


# Tidy the Professor Trash Wheel dataset
prof_trash_df = readxl::read_excel("./p8105_hw2_data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                                 sheet = "Professor Trash Wheel",
                                 range = cellranger::cell_cols("A:N")) %>% 
  janitor::clean_names() %>%
  drop_na(dumpster) %>% 
  filter(dumpster != "Grand Total") %>%
  mutate(sports_balls = as.integer(round(sports_balls, 0))) %>% 
  rename(weight = weight_tons, volume = volume_cubic_yards)


# Merge the two Trash Wheel datasets
mr_trash_merge_df = mr_trash_df %>% mutate(source = "Mr. Trash Wheel")
prof_trash_merge_df = prof_trash_df %>% mutate(source = "Professor Trash Wheel")
trash_merge_df = rbind(mr_trash_merge_df, prof_trash_merge_df)


# Total weight of trash collected by Professor Trash Wheel
total_weight_prof = sum(prof_trash_df$weight)


# Total number of sports balls collected by Mr. Trash Wheel in 2020
total_ball_mr = mr_trash_df %>% 
  filter(year == "2020") %>% 
  pull(sports_balls) %>% 
  sum()
```

* There are `r nrow(trash_merge_df)` observations in the resulting dataset, including `r ncol(trash_merge_df)` variables, namely `r names(trash_merge_df)`.
* The total weight of trash collected by Professor Trash Wheel is `r total_weight_prof` tons.
* The total number of sports balls collected by Mr. Trash Wheel in 2020 was `r total_ball_mr`.

&nbsp;


## Problem 3 (Fivethirtyeight datasets)
```{r}
# Tidy the pols_month dataset
pols_df = read.csv("./p8105_hw2_data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, c("year","month","day")) %>%
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day))%>%
  mutate(month = month.abb[month]) %>% 
  mutate(president = ifelse(prez_gop == "0", "dem","gop")) %>% 
  relocate(president, .before = "gov_gop") %>% 
  select(-prez_dem, -prez_gop, -day)


# Tidy the snp dataset
snp_df = read.csv("./p8105_hw2_data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::parse_date_time2(date,orders ="mdy", cutoff_2000 = 23)) %>% 
  separate(date, c("year","month","day")) %>%
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day))%>%
  mutate(month = month.abb[month]) %>% 
  relocate(year, .before = "month") %>% 
  select(-day)


# Tidy the unemployment dataset
unemployment_df = read.csv("./p8105_hw2_data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(Jan:Dec,names_to = "month", values_to = "unemployment_rate") %>%
  janitor::clean_names() 


# Merge snp into pols
snp_into_pols =  left_join(pols_df, snp_df, by = c("year","month"))


# Merge unemployment into the result
merge_df = left_join(snp_into_pols, unemployment_df, by = c("year","month"))
```


* All of the three datasets used in this problem were taken from the `FiveThirtyEight` data.

* The tidied `pols-month` dataset contains `r nrow(pols_df)` observations with `r ncol(pols_df)` variables, which are `r names(pols_df)`, respectively. The range of years is [`r pols_df %>% pull(year) %>% range()`].

* The tidied `snp` dataset contains `r nrow(snp_df)` observations with `r ncol(snp_df)` variables,  which are `r names(snp_df)`, respectively. The range of years is [`r snp_df %>% pull(year) %>% range()`].

* The tidied `unemployment` dataset contains `r nrow(unemployment_df)` observations of `r ncol(unemployment_df)` variables, which are `r names(unemployment_df)`, respectively. The range of years is [`r unemployment_df %>% pull(year) %>% range()`].

* After merging `snp` dataset into `pols-month` dataset, the dimension of the resulting dataset is `r nrow(snp_into_pols)` rows x `r ncol(snp_into_pols)` columns. It includes `r ncol(snp_into_pols)` variables: `r names(snp_into_pols)`. The range of years is [`r snp_into_pols %>% pull(year) %>% range()`].

* After merging `unemployment` dataset into the previous resulting dataset `snp_into_pols`, the dimension of the merged dataset is `r nrow(merge_df)` rows x `r ncol(merge_df)` columns. It includes `r ncol(merge_df)` variables: `r names(merge_df)`. The range of years is [`r merge_df %>% pull(year) %>% range()`].












